---
title: "Regularization of Bayesian NN"
subtitle: "Proposal"
author: 
  - name: "Dropout Dynamics"
    affiliations:
      - name: "College of Information Science, University of Arizona"
description: "This project aims to predict employee stress levels (low, medium, or high) using a neural network enhanced with Monte Carlo Dropout regularization. By applying dropout during both training and inference, we aim to improve the model's generalization and estimate prediction uncertainty. The model will be trained on a diverse dataset of 5,000 employee samples, incorporating factors like age, job role, hours worked, and mental health conditions. The project explores how effectively a neural network can predict stress based on employee metadata and the impact of dropout regularization on model performance, ultimately helping organizations better support employee well-being."
format:
  html:
    code-tools: true
    code-overflow: wrap
    code-line-numbers: true
    embed-resources: true
editor: visual
code-annotations: hover
execute:
  warning: false
---

```{r}
#| label: load-packages
#| include: false

# Load packages here
pacman::p_load(tidyverse)
```

## Dataset

```{r}
#| label: load-dataset
#| message: false

```

A brief description of your dataset including its provenance, dimensions, etc. as well as the reason why you chose this dataset.

Make sure to load the data and use inline code for some of this information.

## Questions

The two questions you want to answer.

## Analysis plan

-   A plan for answering each of the questions including the variables involved, variables to be created (if any), external data to be merged in (if any).
